Fast approximate Bayesian inference for small-area estimation of HIV indicators using the Naomi model

### 2023-05-24 src/docs_paper/paper.Rmd
Add KS results for latent field to manuscript

The two-sample Kolmogorov-Smirnov (KS) test statistic [@smirnov1948table] is the maximum absolute difference between two empirical cumulative distribution (ECDF) functions $F(\varphi) = \frac{1}{n} \sum_{i = 1}^n \mathbb{I}_{\phi_i \leq \varphi}$.
We compare the KS statistics
\begin{align*}
D_\text{TMB} &= \sup_\varphi | F_\text{NUTS}(\varphi) - F_\text{TMB}(\varphi)|, \\
D_\text{PCA-AGHQ} &= \sup_\varphi | F_\text{NUTS}(\varphi) - F_\text{PCA-AGHQ}(\varphi)|.
\end{align*}
See a summary of the results in Table and Figure \ref{fig:ks}, and full results available in the appendix.

```{r ks, fig.cap="Results of Kolmogorov-Smironv tests analysis."}
qplot(1:3, 1:3) + theme_minimal()
```

```{r}
ks_summary <- readRDS("depends/ks_summary_out.rds")

# gt::gt(ks_summary) %>% gt::as_latex()
```

@@

As well as marginal distributions, we are interested in assessing the accuracy of joint distributions.
Let $\{\bphi_i\}_{i = 1}^n$ be generic samples from a joint posterior.
Pareto-smoothed importance sampling (PSIS) [@vehtari2015pareto, @yao2018yes] is a method for stabilising the ratios $r_s = p(\theta_s, y) / q(\theta)$ used in importance sampling.
Let
\begin{equation*}
p(y \, | \, \mu, \sigma, k) =
\begin{cases}
\frac{1}{\sigma} \left( 1 + k \left( \frac{y - \mu}{\sigma} \right) \right)^{- \frac{1}{k} - 1}, \quad k \neq 0 \\
\frac{1}{\sigma} \exp \left( \frac{y - \mu}{\sigma} \right), \quad k = 0
\end{cases}
\end{equation*}
be the generalised Pareto distribution with shape parameter $k$ and location-scale parameter $(\mu, \tau)$.
See a summary of the results in Table and Figure \ref{fig:psis}.

>>>>

The two-sample Kolmogorov-Smirnov (KS) test statistic [@smirnov1948table] is the maximum absolute difference between two empirical cumulative distribution (ECDF) functions $F(\varphi) = \frac{1}{n} \sum_{i = 1}^n \mathbb{I}_{\phi_i \leq \varphi}$.
We compare the KS statistics
\begin{align*}
D_\text{TMB} &= \sup_\varphi | F_\text{NUTS}(\varphi) - F_\text{TMB}(\varphi)|, \\
D_\text{PCA-AGHQ} &= \sup_\varphi | F_\text{NUTS}(\varphi) - F_\text{PCA-AGHQ}(\varphi)|.
\end{align*}
See a summary of the results in Table.

```{r ks, fig.cap="Results of Kolmogorov-Smironv tests analysis."}
ks_summary <- readRDS("depends/ks_summary.rds")

ks_summary %>%
  filter(Type == "Latent field") %>%
  group_by(Parameter) %>%
  summarise(
    `D(TMB)` = round(mean(`KS(TMB, tmbstan)`), 3),
    `D(PCA-AGHQ)` = round(mean(`KS(aghq, tmbstan)`), 3)
  ) %>%
  gt::gt(rowname_col = "Parameter") %>%
  gt::summary_rows(
   columns = everything(),
   fns = list(Average = ~ mean(.x)),
   formatter = gt::fmt_number,
   decimals = 3
  )
```

@@

As well as marginal distributions, we are interested in assessing the accuracy of joint distributions.
Let $\{\bphi_i\}_{i = 1}^n$ be generic samples from a joint posterior.
Pareto-smoothed importance sampling (PSIS) [@vehtari2015pareto, @yao2018yes] is a method for stabilising the ratios used in importance sampling.
See a summary of the results in Table and Figure \ref{fig:psis}.
